---
title: "SQL_R_TWO_ways"
author: "S.H.Pozveh"
date: "26/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sqldf)
library(odbc)
library(DBI)
library(RSQLite)

```


# 1- SQLDF

- "sqldf" is a package allowing us to write SQL inquries inside R. 

- here I use NHANES datset and NHANES stands for National Health and Nutritional Examination Survue. The data can be donwnloaded using a package with similar name

using "sqldf" is very easy below you could see some of the example.

```{r}
library(NHANES)

## NHANES dataset
nh <- NHANES::NHANES

# before going through the details I am going to split my NHANES datasets to two datasets so I can join them later 
nh1 <- nh %>%
  select(ID, Gender, Age, Education)

nh2 <- nh %>%
  select(ID,Poverty)

## select all the columns of nh where age is abov 50 
sqldf(x = 
        "SELECT * FROM nh
        WHERE Age > 50
      ")

# select the required columns and change their name. 

sqldf(x = 
        "SELECT Gender as Sex, Age, Poverty as 'Income Ratio' FROM nh
      ")
```


## DBPLYR

- dplyr is a pacakge that is written on top of SQL. However, dplyr has an important advantange that is we can use the piping (%>%) while we are writing our queries. Therefore, we will be able to produce rather complicated quriest and if needed we can transform it to SQL queries. 

- Transforming between dplyr and SQL one needs to use a package called dbplyr. dbplyr is in fact is used for writing dplyr-type queires within inside databases.  

- First we are required to create a connection with a database and transfer the dataset to the connection. 

```{r}
# create a SQLite connection insde the memeory
con <- dbConnect(SQLite(), ":memory:")

# copy the nh datasets inside the connection

copy_to(con,nh)
copy_to(con,nh1)
copy_to(con,nh2)

 
```

we can also use dbGetquery() to write SQL queries 

```{r}
dbGetQuery(con,
           'SELECT AVG(Age) as `Mean Age`, Gender
           FROM nh
           GROUP BY Gender'
           )
```

so how does the dbplyr work? 



```{r}
nh
# work with nh dataset
# tell R what dataset we are working on 
nh_db <- tbl(con, "nh")



nh_query <- nh_db %>%
  select(Gender,Age,Education, Poverty , BMI) %>%
  filter(Age>19, 
         Gender == "Male"
         ) %>%
  group_by(Education) %>%
  summarise(mean_age = mean(Age,na.rm=T),
            median_body_mass  = median(BMI,na.rm=T)
            )
# then we can use show_query() function to transform dplyr query to SQL querty
nh_query %>%
  show_query()


## here is the SQL query
# SELECT `Education`, AVG(`Age`) AS `mean_age`, MEDIAN(`BMI`) AS `median_body_mass`
# FROM (SELECT `Gender`, `Age`, `Education`, `Poverty`, `BMI`
# FROM `nh`)
# WHERE ((`Age` > 19.0) AND (`Gender` = 'Male'))
# GROUP BY `Education`
```



## Joining 

```{r}


nh1_db <- tbl(con, "nh1")
nh2_db <- tbl(con, "nh2")

join_query <- nh1_db %>%
  select(ID, Gender, Age) %>%
  filter(Age>30) %>%
  mutate(age_sqr=sqr(Age)) %>%
  mutate(age_mutate = case_when(
    age > 40 ~ sqrt(Age), T ~ Age
  )) %>%
  left_join(nh2_db, by="ID") %>%
  group_by(Gender) %>%
  summarise(across(c(Age, Poverty),mean,na.rm=T,.names = "{.col}_mean"))
  

join_query %>%
  show_query()


# here is the results
# SELECT `Gender`, AVG(`Age`) AS `Age_mean`, AVG(`Poverty`) AS `Poverty_mean`
# FROM (SELECT `LHS`.`ID` AS `ID`, `Gender`, `Age`, `age_sqr`, `age_mutate`, `Poverty`
# FROM (SELECT `ID`, `Gender`, `Age`, `age_sqr`, CASE
# WHEN (`age` > 40.0) THEN (SQRT(`Age`))
# WHEN (1) THEN (`Age`)
# END AS `age_mutate`
# FROM (SELECT `ID`, `Gender`, `Age`, sqr(`Age`) AS `age_sqr`
# FROM (SELECT `ID`, `Gender`, `Age`
# FROM `nh1`)
# WHERE (`Age` > 30.0))) AS `LHS`
# LEFT JOIN `nh2` AS `RHS`
# ON (`LHS`.`ID` = `RHS`.`ID`)
# )
# GROUP BY `Gender`
```








